---
title: "EDA of Monica Dataset"
author: "DATAPROS"
output: html_document
---


```{r basicfcn, include=F}
# can add quietly=T option to the require() function
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
```

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
monica <- data.frame(read.csv("D:/GWU/INRO TO DATA SCIENCE/PROJECT/monica.csv", header = TRUE, stringsAsFactors = F))
library(dplyr)
library(ggplot2)
library(plyr)
library(corrplot)
```

# Inspect dataset: Data Dictionary
```{r 1, echo=FALSE}
names(monica)
#head(monica, 10)
#tail(monica, 10)
str(monica)
library(reshape2)
monica_pivot <- melt(monica, id.vars = c("outcome", "sex","age","yronset"))
#head(monica_pivot)
```

Dataframe with 13 variables (including X, counter of number of samples) and 6367 samples.

Column 1 labeled "3" contains a counter of the samples.

12 variables to analyze: 3 are numeric - age and yronset - and the data type of the remaining 10 is characters.


**Inspect the 10 columns with data type characters**

```{r 2, echo=FALSE}
print('outcome')
table(monica$outcome)
print('sex')
table(monica$sex)
print('premi')
table(monica$premi)
print('smstat')
table(monica$smstat)
print('diabetes')
table(monica$diabetes)
print('highbp')
table(monica$highbp)
print('hicol')
table(monica$hichol)
print('angina')
table(monica$angina)
print('stroke')
table(monica$stroke)
print('hosp')
table(monica$hosp)

```
Large number of Not Known ("nk") values in 7 columns: "premi","smstat","diabetes","highbp","hichol","angina" and "stroke"

3 Columns with complete info: "outcome", "sex" and "hosp"


**Inspect the 2 columns with data type numeric**
```{r 3, echo=FALSE}
#sum(is.na(monica$age))
#sum(is.na(monica$yronset))
#sum(is.null(monica$age))
#sum(is.null(monica$yronset))
summary(monica$age)
summary(monica$yronset)
#describe(monica$age)
```


**Drop column 1**

Drop redundant column (column # 1).
Check operation.

```{r 4, echo=FALSE}
monica <- monica[, 2:13]
str(monica)
```
```{r, echo=FALSE}
#plot by groups of ages and sex
ggplot(data=monica, aes(age, fill=sex)) + 
  geom_bar(color= "#01068a", width = 0.5, position = position_dodge()) + 
  labs(subtitle="Summary by Sex and Ages", y="Total", x="Groups of age") 

#plot by groups of year of onset and sex
ggplot(data=monica, aes(yronset, fill=sex)) + 
  geom_bar(color= "#01068a", width = 0.5, position = position_dodge()) + 
  labs( subtitle="Summary by Sex and Year of Onset", y="Total", x="Year of Onset") 
```

# Inspect distributions



```{r 5, echo=FALSE}
ggplot(aes(x = sex),data = monica) +
    geom_bar(fill = c("green","blue")) +
     scale_x_discrete(labels = c("Females", "Males")) +
    ggtitle("Distribution of the sample by sex")
```



The data sample is skewed towards men.


```{r 6, echo=FALSE}
ggplot(aes(x = age),data = monica) +
    geom_histogram(binwidth = 1,col = "black",fill = "cyan") +
    xlim(c(35,69)) +
    ggtitle("Distribution of the sample by age") +
    scale_x_sqrt()
```


Majority of the samples in this dataset correspond to people from 55 to 75 years old.



```{r 7, echo=FALSE}
monica55to75 <- subset(monica, age >= 55)
ggplot(aes(x = monica55to75$sex),data = monica55to75) +
    geom_bar(fill = c("green","blue")) +
     scale_x_discrete(labels = c("Females", "Males")) +
    labs(title= "Distribution of the Sampleby Sex, Patients 55+", x="Sex", y="Count")
```



The data of samples from 55 to 75 yeard old is skewed towards men as well.

```{r 8, echo=FALSE}
ggplot(aes(x = outcome),data = monica) +
    geom_bar(fill = c("black","orange")) +
     scale_x_discrete(labels = c("Dead", "Live")) +
    ggtitle("Distribution of the sample by outcome")
```



```{r 9, echo=FALSE}
ggplot(aes(x = hosp),data = monica) +
    geom_bar(fill = c("white","purple")) +
     scale_x_discrete(labels = c("Hospotalized", "Not hospitalized")) +
    ggtitle("Distribution of the sample by hospitalization")
```




```{r 10, echo=FALSE}
ggplot(aes(x =smstat),data = monica) +
    geom_bar(fill = c("blue","purple","green","red")) +
     scale_x_discrete(labels = c("Ex-smoker", "No-smoker", "Current-smoker","Not Known")) +
    ggtitle("Distribution of the sample by smoking status")
```





```{r 11, echo=FALSE}
ggplot(aes(x =premi),data = monica) +
    geom_bar(fill = c("blue","coral","cyan")) +
     scale_x_discrete(labels = c("Yes","No","Not Known")) +
    ggtitle("Distribution of the sample by previous miocardial infraction")
```





```{r 12, echo=FALSE}
ggplot(aes(x =diabetes),data = monica) +
    geom_bar(fill = c("darkgreen","green","darkorchid1")) +
     scale_x_discrete(labels = c("Yes","No","Not Known")) +
    ggtitle("Distribution of the sample by diabetes")
```






```{r 13, echo=FALSE}
ggplot(aes(x =highbp),data = monica) +
    geom_bar(fill = c("deepskyblue","darksalmon","green")) +
     scale_x_discrete(labels = c("Yes","No","Not Known")) +
    ggtitle("Distribution of the sample by high blood pressure")
```





```{r 14, echo=FALSE}
ggplot(aes(x =hichol),data = monica) +
    geom_bar(fill = c("deeppink","green","red")) +
     scale_x_discrete(labels = c("Yes","No","Not Known")) +
    ggtitle("Distribution of the sample by high cholestrol")
```







```{r 15, echo=FALSE}
ggplot(aes(x =angina),data = monica) +
    geom_bar(fill = c("purple","aquamarine","deeppink4")) +
     scale_x_discrete(labels = c("Yes","No","Not Known")) +
    ggtitle("Distribution of the sample by angina")
```










```{r 16, echo=FALSE}
ggplot(aes(x =stroke),data = monica) +
    geom_bar(fill = c("darkolivegreen1","green","red")) +
     scale_x_discrete(labels = c("No","Yes","Not Known")) +
    ggtitle("Distribution of the sample by stroke")
```







**Summary**

The variables with complete information are: Sex, Age, Outcome, Hospitalization, and Year of Onset.


# Univariate analysis

**Prepare Dataset for Chi-Square**
```{r 17, echo=FALSE}
#Copy dataset
monica.na <- monica
#Convert all nk in the dataset to NA
monica.na <- na_if(monica.na, 'nk')
#check conversion
str(monica.na)
table(monica.na$premi)
table(monica.na$smstat)
table(monica.na$diabetes)
table(monica.na$highbp)
table(monica.na$hichol)
table(monica.na$angina)
table(monica.na$stroke)
```








**Chi-Square for Each Categorical Variable by Outcome**



This variable was not modified, was already complete.
p-value should be 0.9997 amd it is.





**Sex and Outcome**


```{r 18, echo=FALSE}
#Create table
table11 <- table(monica.na$sex, monica.na$outcome)
# Next perform chi-square test
chi.res11 <- chisq.test(table11)
chi.res11
mosaicplot(~ sex + outcome, data = monica.na,main = "SEX VS OUTCOME", shade = TRUE)


corrplot(chi.res11$residuals, is.cor = FALSE)
```



The null hypothesis H0 is Sex and Outcome are independent.

p-value is greater than 0.05. 

Therefore, we fail to reject the null hypothesis.







**Previous Myocardial Infarction event and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 0.001374. We get the same.

```{r 19, echo=FALSE}
#Create table
table12 <- table(monica.na$premi, monica.na$outcome)
# Next perform chi-square test
chi.res12 <- chisq.test(table12)
chi.res12
mosaicplot(~ premi + outcome, data = monica.na,main = "PREMI VS OUTCOME", shade = TRUE, expected= monica.na$premi)


corrplot(chi.res12$residuals, is.cor = FALSE)
```


The null hypothesis H0 is Previous Myocardial Infarction event and Outcome are independent.

p-value is smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Previous Myocardial Infarction event and Outcome are not significantly independent.




**Smoking status and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 0.03172. We get the same.

```{r ,echo=FALSE}
#Create table
table13 <- table(monica.na$smstat, monica.na$outcome)
# Next perform chi-square test
chi.res13 <- chisq.test(table13)
chi.res13
mosaicplot(~ smstat + outcome, data = monica.na,main = "SMOKING STATUS VS OUTCOME", shade = TRUE, expected= monica.na$smstat)

corrplot(chi.res13$residuals, is.cor = FALSE)
```



The null hypothesis H0 is Smoking status and Outcome are independent.

p-value is smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Smoking status and Outcome are not significantly independent.




**Diabetes and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 5.472e-05. We get the same.

```{r , echo=FALSE}
#Create table
table14 <- table(monica.na$diabetes, monica.na$outcome)
# Next perform chi-square test
chi.res14 <- chisq.test(table14)
chi.res14
mosaicplot(~ diabetes + outcome, data = monica.na,main = "DIABETES VS OUTCOME", shade = TRUE, expected= monica.na$diabetes)

corrplot(chi.res14$residuals, is.cor = FALSE)
```


The null hypothesis H0 is Diabetes and Outcome are independent.

p-value is significantly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Diabetes and Outcome are not significantly independent.




**High Blood Pressure and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 0.1915. We get the same.


```{r ,echo=FALSE}
#Create table
table15 <- table(monica.na$highbp, monica.na$outcome)
# Next perform chi-square test
chi.res15 <- chisq.test(table15)
chi.res15
mosaicplot(~ highbp + outcome, data = monica.na,main = "HIGH BLOOD PRESSURE VS OUTCOME", shade = TRUE, expected= monica.na$highbp)

corrplot(chi.res15$residuals, is.cor = FALSE)
```


The null hypothesis H0 is High Blood Pressure and Outcome are independent.

p-value is gretaer than 0.05.

Therefore, we fail to reject the null hypothesis H0.




**High Cholesterol and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 1.182e-10. We get the same.

```{r , echo=FALSE}
#Create table
table16 <- table(monica.na$hichol, monica.na$outcome)
# Next perform chi-square test
chi.res16 <- chisq.test(table16)
chi.res16
mosaicplot(~ hichol + outcome, data = monica.na,main = "HIGH CHOLESTROL VS OUTCOME", shade = TRUE, expected= monica.na$hichol)

corrplot(chi.res16$residuals, is.cor = FALSE)
```


The null hypothesis H0 is High Cholesterol and Outcome are independent.

p-value is significantly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that High Cholesterol and Outcome are not significantly independent.





**Angina and Outcome**



This variable was modified, nk replaced with NA.
p-value of subset is 2.137e-07. We get the same.

```{r , echo=FALSE}
#Create table
table17 <- table(monica.na$angina, monica.na$outcome)
# Next perform chi-square test
chi.res17 <- chisq.test(table17)
chi.res17

mosaicplot(~ angina + outcome, data = monica.na,main = "ANGINA VS OUTCOME", shade = TRUE, expected= monica.na$angina)

corrplot(chi.res17$residuals, is.cor = FALSE)
```



The null hypothesis H0 is Angina and Outcome are independent.

p-value is significantly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Angina and Outcome are not significantly independent.






**Stroke and Outcome**


This variable was modified, nk replaced with NA.
p-value of subset is 3.897e-13. We get the same.

```{r , echo=FALSE}
#Create table
table18 <- table(monica.na$stroke, monica.na$outcome)
# Next perform chi-square test
chi.res18 <- chisq.test(table18)
chi.res18

mosaicplot(~ stroke + outcome, data = monica.na,main = "STROKE VS OUTCOME", shade = TRUE, expected= monica.na$stroke)

corrplot(chi.res18$residuals, is.cor = FALSE)
```



The null hypothesis H0 is Stroke and Outcome are independent.

p-value is significantly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Stroke and Outcome are not significantly independent.






**Hospitalization and Outcome**



This variable was not modified, was already complete.
p-value should be < 2.2e-16 amd it is.

```{r , echo=FALSE}
#Create table, no need to subset
table19 <- table(monica.na$hosp, monica.na$outcome)
# Next perform chi-square test
chi.res19 <- chisq.test(table19)
chi.res19

mosaicplot(~ hosp + outcome, data = monica.na,main = "HOSPITAL VS OUTCOME", shade = TRUE, expected= monica.na$hosp)

corrplot(chi.res19$residuals, is.cor = FALSE)
```



The null hypothesis H0 is Hospitalization and Outcome are independent.

p-value is significanly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Hospitalization and Outcome are not significantly independent.




# Summary of univariate analysis

Variables for which we reject the null hypothesis Ho and conclude that they are not significantly independent:
Previous Myocardial Infarction, Smoking Status, Diabetes, High Cholesterol, Angina, Stroke and Hospitalization.

Among these variables, p-value is significantly smaller for: Diabetes, High Cholesterol, Angina, Stroke and Hospitalization.

Variables for which we fail to reject the null hypothesis:
Sex and High Blood Pressure.


**Summary of Risk Factors**

| Risk factors    | Reject/Fail to reject | p-value |
|-----------------|-----------------------|---------|
|- myocardial inf.|Reject                |`r chi.res12$p.value`|
| - smoking status |Reject               |`r round(chi.res13$p.value, digits=5)`|
| - diabetes       |Reject               |`r round(chi.res14$p.value, digits=5)` |
|- high blood pressure |Fail to reject   |`r chi.res15$p.value`  |
| - high colesterol|Reject               |`r chi.res16$p.value`  |
| - angina         |Reject               |`r chi.res17$p.value` |
| - stroke         |Reject               |`r chi.res18$p.value`  |
| - hospitalization|Reject               |`r chi.res19$p.value`  |


# Univariate analysis stratified by Sex


```{r 20, echo=FALSE}
#Subset by sex
monica.f <- subset(monica.na, sex == 'f')
monica.m <- subset(monica.na, sex == 'm')
str(monica.f)
str(monica.m)
```


**Sex and Outcome by Sex**


Female

```{r 22.1, echo=FALSE}
#Create table
table22.1 <- table(monica.f$sex, monica.f$outcome)
# Next perform chi-square test
chi.res22.1 <- chisq.test(table22.1)
chi.res22.1

#corrplot(chi.res22.1$residuals, is.cor = FALSE)
```


Male


```{r 22.2, echo=FALSE}
#Create table
table22.2 <- table(monica.m$sex, monica.m$outcome)
# Next perform chi-square test
chi.res22.2 <- chisq.test(table22.2)
chi.res22.2


#corrplot(chi.res22.2$residuals, is.cor = FALSE)
```



Same conclusion for female and male.




**Previous Myocardial Infarction and Outcome by Sex**


Female



```{r 23.1, echo=FALSE}
#Create table
table23.1 <- table(monica.f$premi, monica.f$outcome)
# Next perform chi-square test
chi.res23.1 <- chisq.test(table23.1)
chi.res23.1

corrplot(chi.res23.1$residuals, is.cor = FALSE)
```


Male



```{r 23.2, echo=FALSE}
#Create table
table23.2 <- table(monica.m$premi, monica.m$outcome)
# Next perform chi-square test
chi.res23.2 <- chisq.test(table23.2)
chi.res23.2

corrplot(chi.res23.2$residuals, is.cor = FALSE)
```



Different conclusion for female and male!




**Smoking Status and Outcome by Sex**


Female



```{r 24.1, echo=FALSE}
#Create table
table24.1 <- table(monica.f$smstat, monica.f$outcome)
# Next perform chi-square test
chi.res24.1 <- chisq.test(table24.1)
chi.res24.1

corrplot(chi.res24.1$residuals, is.cor = FALSE)
```



Male



```{r 24.2, echo=FALSE}
#Create table
table24.2 <- table(monica.m$smstat, monica.m$outcome)
# Next perform chi-square test
chi.res24.2 <- chisq.test(table24.2)
chi.res24.2

corrplot(chi.res24.2$residuals, is.cor = FALSE)
```



Different conclusion for female and male!



**Diabetes and Outcome by Sex**


Female


```{r 25.1, echo=FALSE}
#Create table
table25.1 <- table(monica.f$diabetes, monica.f$outcome)
# Next perform chi-square test
chi.res25.1 <- chisq.test(table25.1)
chi.res25.1

corrplot(chi.res25.1$residuals, is.cor = FALSE)
```


Male


```{r 25.2, echo=FALSE}
#Create table
table25.2 <- table(monica.m$diabetes, monica.m$outcome)
# Next perform chi-square test
chi.res25.2 <- chisq.test(table25.2)
chi.res25.2

corrplot(chi.res25.2$residuals, is.cor = FALSE)
```



Same conclusion for female and male.



**High Blood Pressure and Outcome by Sex**


Female


```{r 26.1, echo=FALSE}
#Create table
table26.1 <- table(monica.f$highbp, monica.f$outcome)
# Next perform chi-square test
chi.res26.1 <- chisq.test(table26.1)
chi.res26.1

corrplot(chi.res26.1$residuals, is.cor = FALSE)
```



Male


```{r 26.2, echo=FALSE}
#Create table
table26.2 <- table(monica.m$highbp, monica.m$outcome)
# Next perform chi-square test
chi.res26.2 <- chisq.test(table26.2)
chi.res26.2

corrplot(chi.res26.2$residuals, is.cor = FALSE)
```


Same conclusion for female and male.




**High Cholesterol and Outcome by Sex**


Female


```{r 27.1, echo=FALSE}
#Create table
table27.1 <- table(monica.f$hichol, monica.f$outcome)
# Next perform chi-square test
chi.res27.1 <- chisq.test(table27.1)
chi.res27.1

corrplot(chi.res27.1$residuals, is.cor = FALSE)
```


Male



```{r 27.2, echo=FALSE}
#Create table
table27.2 <- table(monica.m$hichol, monica.m$outcome)
# Next perform chi-square test
chi.res27.2 <- chisq.test(table27.2)
chi.res27.2

corrplot(chi.res27.2$residuals, is.cor = FALSE)
```


Same conclusion for female and male.


**Angina and Outcome by Sex**


Female


```{r 28.1, echo=FALSE}
#Create table
table28.1 <- table(monica.f$angina, monica.f$outcome)
# Next perform chi-square test
chi.res28.1 <- chisq.test(table28.1)
chi.res28.1

corrplot(chi.res28.1$residuals, is.cor = FALSE)
```


Male



```{r 28.2, echo=FALSE}
#Create table
table28.2 <- table(monica.m$angina, monica.m$outcome)
# Next perform chi-square test
chi.res28.2 <- chisq.test(table28.2)
chi.res28.2

corrplot(chi.res28.2$residuals, is.cor = FALSE)
```



Different conclusion for female and male!



**Stroke and Outcome by Sex**



Female



```{r 29.1, echo=FALSE}
#Create table
table29.1 <- table(monica.f$stroke, monica.f$outcome)
# Next perform chi-square test
chi.res29.1 <- chisq.test(table29.1)
chi.res29.1

corrplot(chi.res29.1$residuals, is.cor = FALSE)
```


Male


```{r 29.2, echo=FALSE}
#Create table
table29.2 <- table(monica.m$stroke, monica.m$outcome)
# Next perform chi-square test
chi.res29.2 <- chisq.test(table29.2)
chi.res29.2

corrplot(chi.res29.2$residuals, is.cor = FALSE)
```


same conclusion for female and male.



**Hospitalization and Outcome by Sex**



Female




```{r 30.1, echo=FALSE}
#Create table
table30.1 <- table(monica.f$hosp, monica.f$outcome)
# Next perform chi-square test
chi.res30.1 <- chisq.test(table30.1)
chi.res30.1


corrplot(chi.res30.1$residuals, is.cor = FALSE)
```


Male



```{r 30.2, echo=FALSE}
#Create table
table30.2 <- table(monica.m$hosp, monica.m$outcome)
# Next perform chi-square test
chi.res30.2 <- chisq.test(table30.2)
chi.res30.2

corrplot(chi.res30.2$residuals, is.cor = FALSE)
```


Same conclusion for female and male.


**Summary Female vs Risk factors**

| Risk factor | Reject/Fail to reject | p-value     |
|-----------------|-------------------| ------------|
|- myocardial inf. |Reject            |`r round(chi.res23.1$p.value, digits=4)` |
| - smoking status |Reject            |`r chi.res24.1$p.value`|
| - diabetes       |Fail to reject    |`r chi.res25.1$p.value`|
|- high blood pressure |Reject        |`r chi.res26.1$p.value`|
| - high colesterol|Reject            |`r chi.res27.1$p.value`|
| - angina         |Reject            |`r chi.res28.1$p.value`|
|- stroke          |Reject            |`r chi.res29.1$p.value`|
|- sex             |Reject            |`r chi.res22.1$p.value`|
|- hospitalization |Reject            |`r chi.res30.1$p.value`|


**Summary Male vs Risk factors**

| Risk factor | Reject/Fail to reject | p-value     |
|-----------------|-------------------| ------------|
|- myocardial inf. |Fail to Reject    |`r chi.res23.2$p.value` |
| - smoking status |Fail to Reject    |`r chi.res24.2$p.value`|
| - diabetes       |Reject            |`r chi.res25.2$p.value`|
|- high blood pressure |Fail to Reject|`r chi.res26.2$p.value`|
| - high colesterol|Reject            |`r chi.res27.2$p.value` |
| - angina         |Fail to Reject    |`r chi.res28.2$p.value`|
|- stroke          |Reject            |`r chi.res29.2$p.value`|
|- sex             |Reject            |`r chi.res22.2$p.value`|
|- hosppitalization|Reject            |`r chi.res30.2$p.value`|

**Study of Missingness of each Feature by Sex**

```{r 31.1, echo=FALSE}
#Subset dataset by sex
monica.fnk <- subset(monica, monica$sex == 'f')
monica.mnk <- subset(monica, monica$sex == 'm')
#premi
print('premi / female')
table(monica.fnk$premi)
print('premi / male')
table(monica.mnk$premi)
#smstat
print('smstat / female')
table(monica.fnk$smstat)
print('smstat / male')
table(monica.mnk$smstat)
#diabetes
print('diabetes / female')
table(monica.fnk$diabetes)
print('diabetes / male')
table(monica.mnk$diabetes)
#highbp
print('highbp / female')
table(monica.fnk$highbp)
print('highbp / male')
table(monica.mnk$highbp)
#hichol
print('hichol / female')
table(monica.fnk$hichol)
print('hichol / male')
table(monica.mnk$hichol)
#angina
print('angina / female')
table(monica.fnk$angina)
print('angina / male')
table(monica.mnk$angina)
#stroke
print('stroke / female')
table(monica.fnk$stroke)
print('stroke / male')
table(monica.mnk$stroke)
```

# Univariate analysis of numerical variables Age and Year of Onset

**Prepare Dataframe**

```{r 32.1, echo=FALSE}
#copy dataframe
monica.na.bin <- monica.na
#inspect age
summary(monica.na.bin$age)
#bin age in new column
monica.na.bin$agecat <- cut(monica.na.bin$age, c(34.5, 39.5, 44.5, 49.5, 54.5, 59.5, 64.5, 69.5))
summary(monica.na.bin$agecat)
str(monica.na.bin)
```

```{r 32.2, echo=FALSE}
#inspect yronset
summary(monica.na.bin$yronset)
#bin yronset in new column
monica.na.bin$yronsetcat <- cut(monica.na.bin$yronset, c(8.45, 89.5, 94.5))
summary(monica.na.bin$yronsetcat)
str(monica.na.bin)
```
```{r 32.3, echo=FALSE}

#Subset by sex
monica.na.bin.f <- subset(monica.na.bin, sex == 'f')
monica.na.bin.m <- subset(monica.na.bin, sex == 'm')
str(monica.na.bin.f)
str(monica.na.bin.m)
```

**Analysis of Age / Outcome and Age / Outcome Stratified by Sex**


**Age / Outcome**
```{r 33, echo=FALSE}
#Create table
table33 <- table(monica.na.bin$agecat, monica.na.bin$outcome)
# Next perform chi-square test
chi.res33 <- chisq.test(table33)
chi.res33
```
The null hypothesis H0 is Age and Outcome are independent.

p-value is significanly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Age and Outcome are not significantly independent.

**Age / Outcome Stratified by Sex**


**Female**
```{r 34, echo=FALSE}
#Create table
table34 <- table(monica.na.bin.f$agecat, monica.na.bin.f$outcome)
# Next perform chi-square test
chi.res34 <- chisq.test(table34)
chi.res34
```
The null hypothesis H0 is Age and Outcome are independent.

p-value is significanly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Age and Outcome are not significantly independent.


**Male**
```{r 35, echo=FALSE}
#Create table
table35 <- table(monica.na.bin.m$agecat, monica.na.bin.m$outcome)
# Next perform chi-square test
chi.res35 <- chisq.test(table35)
chi.res35
```
The null hypothesis H0 is Age and Outcome are independent.

p-value is smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Age and Outcome are not significantly independent.


## Analysis of Year of Onset / outcome and Year of Onset / Outcome stratified by Sex


**Year of Onset / Outcome**

```{r 36, echo=FALSE}
#Create table
table36 <- table(monica.na.bin$yronsetcat, monica.na.bin$outcome)
# Next perform chi-square test
chi.res36 <- chisq.test(table36)
chi.res36
```

The null hypothesis H0 is Year of Onset and Outcome are independent.

p-value is significanly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Year of Onset and Outcome are not significantly independent.


**Fenale**
```{r 37, echo=FALSE}
#Create table
table37 <- table(monica.na.bin.f$yronsetcat, monica.na.bin.f$outcome)
# Next perform chi-square test
chi.res37 <- chisq.test(table37)
chi.res37
```
The null hypothesis H0 is Year of Onset and Outcome are independent.

p-value is slightly smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that Year of Onset and Outcome are not significantly independent.


**Male**
```{r 38, echo=FALSE}
#Create table
table38 <- table(monica.na.bin.m$yronsetcat, monica.na.bin.m$outcome)
# Next perform chi-square test
chi.res38 <- chisq.test(table38)
chi.res38
```
The null hypothesis H0 is yronset and outcome are independent.

p-value is smaller than 0.05.

Therefore, we reject the null hypothesis H0 and conclude that yronset and outcome are not significantly independent.

# Univariate losgistic regression analysis (performed only on values with stastical significance):

**Results for Outcome vs Diabetes**
```{r 39, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val1<-glm(outcome ~ diabetes, data = monica.co, family = "binomial")
summary(val1)$coefficients
```
Since the value of the coefficient is a negative value we can say that having Diabetes does not affect the Outcome.

**Results for Outcome vs Stroke**
```{r 40, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val2<-glm(outcome ~ stroke, data = monica.co, family = "binomial")
summary(val2)$coefficients
```
Since the value of the coefficient is a negative value we can say that history of a Stroke does not affect the Outcome.

**Results for Outcome vs Angina**
```{r 41, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val3<-glm(outcome ~ angina, data = monica.co, family = "binomial")
summary(val3)$coefficients
```
Since the value of the coefficient is a negative value we can say that having Angina does not affect the Outcome.

**Results for Outcome vs High Cholesterol**
```{r 42, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val4<-glm(outcome ~ hichol, data = monica.co, family = "binomial")
summary(val4)$coefficients
```
Since the value of the coefficient is a positive value we can say that having High Cholestrol does affect the Outcome.



**Results for Outcome vs Hospitalization**
```{r 44, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val5<-glm(outcome ~ hosp, data = monica.co, family = "binomial")
summary(val5)$coefficients
```
Since the value of the coefficient is a positive value we can say that being Hospitalized does affect the Outcome.

**Results for Outcome vs Previous Myocardial Infraction event**
```{r 45, echo=FALSE}
#copy dataset
monica.co <- monica.na
#create new column and convert outcome to 0 and 1
monica.co$coutcome <- monica$outcome
monica.co$outcome <-  revalue(monica.co$outcome, c("live"=0, "dead"=1))
monica.co$outcome <- as.factor(monica.co$outcome)
#inspect outcome
summary(monica.co$outcome)
# Next perform logistic regression
val6<-glm(outcome ~ premi, data = monica.co, family = "binomial")
summary(val6)$coefficients
```

Since the value of the coefficient is a negative value we can say that having a Previous Myocardial Infraction event does not affect the outcome.